---
title: "PA-Practical Machine Learning / Weight Lifting Exercise Manner Prediction"
author: "A. Alvarez"
date: "Wednesday, April 22, 2015"
output: html_document
---

##Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.
The goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

##Getting and cleaning data
The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
training <- read.csv("pml-training.csv", na.strings=c("", "NA", "NULL"))
testing <- read.csv("pml-testing.csv", na.strings=c("", "NA", "NULL"))

# We remove the rows of the dataset that don't provide valuable information for fiting the model later on:
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]

# We also remove some unnecessary columns which contain id related variables:
training <- training[, c(7:60)]
testing <- testing[, c(7:60)]
```

##Model building

First, we create the data partitions of training and validating data sets. This splitting is 60% for the training and 40% for the testing.
```{r}
library(caret)
library(randomForest)

inTrain <- createDataPartition(y = training$classe, p=0.60, list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]

set.seed(12345)
model <- randomForest(classe~.,data=train)
print(model)
```

##Model Evaluate
We calculate the out of sample accuracy of the model. In other words, this describes how accurately the model performs on the 40% testing dataset.
```{r}
test_pred <- predict(model, test)
confusionMatrix(test_pred, test$classe)

importance(model)
```
And confirmed the accurancy at validating data set by calculate it with the formula:
```{r accuracy}
accuracy<-c(as.numeric(predict(model,newdata=test[,-ncol(test)])==test$classe))
accuracy<-sum(accuracy)*100/nrow(test)
```
Model Accuracy as tested over Validation set = `r accuracy`%.

## Prediciting with Testing Dataset
Using the random forests model built above, we provide answers for the 20 test cases of the testing dataset.
```{r}
pred_testing <- predict(model, testing)
pred_testing <- as.character(pred_testing)
pred_testing
```

## Generating Answers Files to Submit for Assignment
The following function to create the files to answers the Prediction Assignment Submission:
```{r}
write_files <- function(x) {
        n <- length(x)
        for (i in 1:n) {
                filename <- paste0("problem_id", i, ".txt")
                write.table(x[i], file=filename, quote=FALSE, row.names=FALSE,col.names=FALSE)
        }
}
write_files(pred_testing)
```
